---
title: "PRIM"
author: "Bin Han"
date: "April 29, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(data.table)
```


__Introduction__:

This file combines all the data from CourtNet System and identify potential features.

```{r}
data_path = "~/Duke/Cynthia Research/data/ky-data-raw/"
```

### KY1 -- defendants
```{r, message=FALSE, warning=FALSE}
Defendants = as.data.frame(fread(paste0(data_path,'01_KY_Recidivism_Defendants.txt'), 
                                 header = T)) %>% select(PersonID, Gender, Race)

CNet_cases = as.data.frame(fread(paste0(data_path,'03_KY_Recidivism_CNet_Cases.txt'), 
                                 header = T)) %>% filter(County != "WHITLEY")

CNet_charges = as.data.frame(fread(paste0(data_path,'06_KY_Recidivism_CNet_Charges.txt'),
                                   header = T)) %>% 
    rename(PersonID = PersonId) %>% 
    filter(County != "WHITLEY")

Sentences = as.data.frame(fread(paste0(data_path,'08_KY_Recidivism_Sentences.txt'), 
                                header = T)) %>% 
    rename(PersonID = PersonId) %>% 
    filter(County != "WHITLEY")

```

## EDA 

## 1. Defendants
The original data contains variables including PersonID, Gender, Race, Ethnicity, ID Source. Only kept PersonID and Gender. Other features will not be used in prediction.

```{r}
colnames(Defendants)
```

```{r}
summary(Defendants)
```

```{r}
length(unique(Defendants$PersonID)) == nrow(Defendants)
```

__NOTE__:
Each row has an unique PersonID.

###### 2. CNet_Cases ###############################################################################

```{r}
colnames(CNet_cases)
```

__NOTE__:
Potentially useful features: PersonID, SeqCaseID, PRIM, CaseNumberType, FilingDate, Age_at_Filing (CountyCode, County and Division could be used to do sanity check)

```{r}
summary(CNet_cases)
```

```{r}
sum(CNet_cases$PRIM)
```

__NOTE__:
Not all the cases are in the PRIM system. We might want to only work on cases that appear in the PRIM, which helps reduce the data size.

```{r}
length(unique(CNet_cases$PersonID)) == nrow(CNet_cases)
length(unique(CNet_cases$SeqCaseID)) == nrow(CNet_cases)
```

__NOTE__:
Some PersonID's appear multiple times, which means that a PersonID could relate to multiple cases. Each case has an unique SeqCaseID.

__Sanity Check__:
It is confirmed that some people did have more than one cases. 

```{r}
CNet_cases %>% group_by(PersonID) %>% summarize(count = n()) %>% arrange(desc(count))
```

```{r}
CNet_cases[CNet_cases$PersonID == "19143", ]
```


###### 3. CNet_charges ###############################################################################

```{r}
colnames(CNet_charges)
```

__NOTE__:
Potentially useful features: PersonID, SeqCaseID, CaseNumberType, ChargeDate, ChargeEntityNumber, ChargeNumber, UOR Descripiton, UorLevel

```{r}
summary(CNet_charges)
```

__NOTE__:
The chargeEntityNumber has a 7000; suspeciously high.

```{r}
length(unique(CNet_charges$PersonID)) == nrow(CNet_charges)
length(unique(CNet_charges$SeqCaseID)) == nrow(CNet_charges)
```

__NOTE__:
It is expected that the PersonID is not equal to the number of rows of CNet_charges. Each unique case in the CourtNet system could relate to multiple charges.

__Sanity Check__:
It is confirmed that one case could correspond to several charges. There are people who had over thousands of charges under one case....

```{r}
CNet_charges %>% group_by(PersonID, SeqCaseID) %>% summarize(count = n()) %>% arrange(desc(count))
```

```{r}
CNet_charges[CNet_charges$PersonID == "409581" & CNet_charges$SeqCaseID == "1715219", ]
CNet_charges[CNet_charges$PersonID == "349879" & CNet_charges$SeqCaseID == "1372149", ]
CNet_charges[CNet_charges$PersonID == "299055" & CNet_charges$SeqCaseID == "1131637", ]
```


###### 4. Sentences ################################################################################
```{r}
colnames(Sentences)
```

```{r}
summary(Sentences)
```









##### Combine Tables ###############################################################################

During the whole combination process, we will keep as many features as possible, which aims to keep as much information as possible, through renaming features and conducting sanity checking, which also ensures the correctness of combination.


## 1. Combine cases with defendants #######
```{r, message=FALSE}
memory.limit(size=56000)
comb1 = merge(x=Defendants,
              y=CNet_cases,
              by = "PersonID")

nrow(comb1) == nrow(CNet_cases)
```

__NOTE__: no information added/lost. 

```{r}
colnames(comb1)
```

```{r}
comb1 = comb1 %>% 
    rename(Age_at_Case_Filing = Age_at_Filing,
           Case_FilingDate = FilingDate,
           Case_DispositionDate = DispositionDate)
```





## 2. Combine with charges ########
```{r}
comb2 = merge(x = comb1, 
              y=CNet_charges,
              by = c("PersonID", "SeqCaseID")) 

nrow(comb2) == nrow(CNet_charges)
```

__NOTE__: no information added/lost

```{r}
colnames(comb2)
```

__Sanity Check__:
There is no conflict of County, CountyCode, Division, and CaseNumberType between the cases and charges data sets. 

```{r}
sum(comb2$CountyCode.x == comb2$CountyCode.y) == nrow(comb2)
sum(comb2$County.x == comb2$County.y) == nrow(comb2)
sum(comb2$Division.x == comb2$Division.y) == nrow(comb2)
sum(comb2$CaseNumberType.x == comb2$CaseNumberType.y) == nrow(comb2)
```


```{r}
comb2 = comb2 %>% 
  select(-County.y,
         -CountyCode.y,
         -Division.y,
         -CaseNumberType.y) %>% 
  rename(CountyCode = CountyCode.x,
         County = County.x,
         Division = Division.x,
         Case_NumberType = CaseNumberType.x,
         ChargeType = type,
         Charge_DispositionDate = DispositionDate,
         Charge_DispositionCode = DispositionTypeCode,
         Charge_DispositionType = DispositionType,
         UOR_Class = UorClass,
         UOR_Level = UorLevel,
         UOR_Code = "UOR Code",
         UOR_Description = "UOR Description")
```




## 3. Combine with sentences ########

```{r, warning=FALSE, message=FALSE}
memory.limit(size=56000)
comb3 = merge(x=comb2,
              y = Sentences,
              by = c("PersonID", "SeqCaseID", "ChargeEntityNumber", "ChargeNumber"),
              all.x = TRUE) 

nrow(comb3) == nrow(comb2)
```

__NOTE__: extra rows added to comb3

## check SentenceEntityNumber != ChargeEntityNumber
```{r}
a = Sentences[Sentences$PersonID == "62628" & Sentences$SeqCaseID == "2775564", ]
b = comb2[comb2$PersonID == "62628" & comb2$SeqCaseID == "2775564", ]
```

__NOTE__: ChargeNumber != ChargeEntityNumber; ChargeNumber != SentenceEntityNumber; ChargeEntityNumber != SentenceEntityNumber; For each ChargeEntity number, there could be multiple sentenseEntity number corresponded. 

## check unique sentences
```{r}
a = Sentences %>% 
    group_by(PersonID, SeqCaseID, ChargeEntityNumber, ChargeNumber) %>% 
    summarise(count = n()) %>% 
    arrange(desc(count))
b = a %>% filter(count > 1)
sum(b$count - 1)
```

__NOTE__:

767 is exactly we got for the extra records. It is confirmed that for some charges, there might be more than one sentences. 

## clean names
```{r}
colnames(comb3)
```

```{r}
comb3 = comb3 %>% 
  select(-DispositionTypeCode,
         -DispositionType,
         -DispositionDate,
         -CountyCode.y,
         -County.y,
         -Division.y,
         -ChargeDate.y,
         -CaseNumberType,
         -type,
         -"UOR Code",
         -"UOR Description",
         -UorClass,
         -UorLevel) %>% 
  rename(County = County.x,
         CountyCode = CountyCode.x,
         Division = Division.x,
         ChargeDate = ChargeDate.x)

```


### work with final data
### Subset useful features

Based on what we need, we can subset any features we want.

```{r}
CNet_data = comb3
save(CNet_data, file = "CNet.RData")
```





