---
title: "PRIM"
author: "Bin Han"
date: "May 2, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(data.table)
```

### Data path
```{r}
data_path = "~/Duke/Cynthia Research/data/ky-data-raw/"
```




### Load PRIM data

```{r}
## defendants
Defendants = as.data.frame(fread(paste0(data_path,'01_KY_Recidivism_Defendants.txt'), 
                                 header = T)) %>% select(PersonID, Gender)
## interviews
Interviews = as.data.frame(fread(paste0(data_path,'02_KY_Recidivism_Interviews.txt'))) %>% 
    rename(PersonID = PersonId, InterviewID = InterviewId, Interview_Date_Created = DateCreated)

## cases
PRIM_cases = as.data.frame(fread(paste0(data_path,'04_KY_Recidivism_PRIM_Cases.txt'))) %>%
    rename(SeqCaseID = SeqCaseId, 
           Case_Date_Booked = date_booked, 
           Age_at_Case_Booked = Age_at_Booking) %>%
    filter(County != "WHITLEY")


## charges
PRIM_charges = as.data.frame(fread(paste0(data_path,'07_KY_Recidivism_PRIM_Charges.txt'))) %>%
    rename(InterviewID = interviewID, PersonID = PersonId, CaseID = CaseId, 
           Charge_DispositionDate = DispositionDate) %>% 
    filter(County != "WHITLEY")


## bonds
Bonds = as.data.frame(fread(paste0(data_path,'09_KY_Recidivism_Bonds.txt'))) %>% 
    rename(InterviewID = interviewID, PersonID = PersonId, CaseID = CaseId)%>% 
    filter(County != "WHITLEY")

## events
Events = as.data.frame(fread(paste0(data_path,'10_KY_Recidivism_Events.txt'))) %>% 
    rename(InterviewID = InterviewId,PersonID = PersonId, CaseID = Caseid,SeqCaseID = SeqCaseid) %>% 
    filter(EventCounty != "WHITLEY")

## fta
FTA = as.data.frame(fread(paste0(data_path,'11_KY_Recidivism_FTA.txt'))) %>% 
    rename(InterviewID = InterviewId, PersonID = PersonId, CaseID = Caseid, SeqCaseID = SeqCaseid)

## supervision
Supervision = as.data.frame(fread(paste0(data_path,'12_KY_Recidivism_Supervision.txt'))) %>% 
  rename(InterviewID = InterviewId, PersonID = PersonId, CaseID = Caseid)
```





### EDA
##### 1. Defendants ################################################################################

The original data contains PersonID, Gender, Race, and Ethnicity. We only need PersonID and Gender.

```{r}
colnames(Defendants)
```

```{r}
summary(Defendants)
```

```{r}
length(unique(Defendants$PersonID)) == nrow(Defendants)
```

__NOTE__:

Eahc row has an unique PersonID, representing a unique defendant.





##### 2. Interview ################################################################################

```{r}
colnames(Interviews)
```

```{r}
Interviews = Interviews %>% 
    select(-HoldingCountyCode, -County, -RiskLevel, -RiskScore, -HasHolder, -PostedPrior) 
    
```

__NOTE__:

Remove some variables to reduce dimension before joining tables.


```{r}
summary(Interviews)
```

```{r}
length(unique(Interviews$PersonID)) == nrow(Interviews)
Interviews %>% group_by(PersonID) %>% summarise(count = n()) %>% arrange(desc(count))
Interviews %>% group_by(PersonID, InterviewID) %>% summarise(count = n()) %>% arrange(desc(count))
```

__NOTE__:

Each person (PersonID) could correspond to several interviews. But the combination of PersonID and InterviewID is unique.

```{r}
Interviews[Interviews$PersonID == "28566", ]
```







##### 3. Cases ################################################################################

```{r}
colnames(PRIM_cases)
```


```{r}
summary(PRIM_cases)
PRIM_cases[is.na(PRIM_cases$InterviewID), ]
```

```{r}
hist(PRIM_cases$Age_at_Case_Booked)
sort(unique(PRIM_cases$Age_at_Case_Booked))
sum(PRIM_cases$Age_at_Case_Booked >= 80)
```

__NOTE__:
1. There are 3 NA's on InterviewID, CaseID, Age_at_Case_Booked. Same three. Error.
2. Max age_at_case_booked: 237. And there are 6369 people with age at case booked > 80. Error?




```{r}
PRIM_cases = PRIM_cases %>% 
    select(-CountyCode, -AlternativeReleaseType, AlternativeReleaseDate, -Mcr, -McrDate) %>% 
    filter(!is.na(InterviewID))
```

__NOTE__:
Remove some unwanted features to reduce dimension.

```{r}
length(unique(PRIM_cases$CaseID)) == nrow(PRIM_cases)
length(unique(PRIM_cases$SeqCaseID)) == nrow(PRIM_cases)
```

__NOTE__:
Each row is an unique case with CaseID. But SeqCaseID has some repeated appearance. It indicates that each SeqCaseID could relate to several CaseID's.


```{r}
PRIM_cases %>% group_by(PersonID, InterviewID) %>% summarise(count = n()) %>% arrange(desc(count))
```

```{r}
PRIM_cases[PRIM_cases$PersonID == "110149" & PRIM_cases$InterviewID == "686821", ]
```

```{r}
PRIM_cases %>% group_by(PersonID, InterviewID, CaseID) %>% summarise(count = n()) %>% arrange(desc(count))
```

```{r}
PRIM_cases %>% group_by(PersonID, InterviewID, SeqCaseID) %>% summarise(count = n()) %>% arrange(desc(count))
```

```{r}
PRIM_cases[PRIM_cases$PersonID == "409241" & PRIM_cases$InterviewID == "1375331" & PRIM_cases$SeqCaseID == "1719208", ]
```

```{r}
PRIM_cases %>% group_by(PersonID, SeqCaseId) %>% summarise(count = n()) %>% arrange(desc(count))
```

```{r}
e = PRIM_cases[PRIM_cases$PersonID == "152948" & PRIM_cases$SeqCaseId == "913837", ]
```


__Insights__:
1. Same person, same interview, could correspond to the different SeqCaseID.
2. Same person, same interview, same SeqCaseID, could have different CaseIDs.
3. Same person, same SeqCaseID, could correspond to multiple interview.





##### 4. Charges ################################################################################

```{r}
colnames(PRIM_charges)
```

```{r}
summary(PRIM_charges)
```









##### 5.Bonds ################################################################################
```{r}
colnames(Bonds)
```

```{r}
summary(Bonds)
```







###### Combine files ########################################################################

### 1. defendants + interviews
```{r}
comb1 = merge(x=Defendants, y=Interviews, by="PersonID")
nrow(comb1) == nrow(Interviews)
```

__NOTE__:
No information lost or added. 

```{r}
colnames(comb1)
```








### 2. comb1 + PRIM_cases
```{r}
## inner join
comb2 = merge(x=comb1, y=PRIM_cases, by=c("PersonID", "InterviewID"))
nrow(comb2) == nrow(PRIM_cases)
```

__NOTE__:
No information added or lost.


```{r}
colnames(comb2)
```

__Summary__:

1. "comb2" contains all the interview and case information in PRIM, that could be related to the CourtNet cases.

2. 5000+ records have age over 90+. Prob typo.

"comb2" with be continuously merged with other files including "Bonds", "Events", "FTA" etc, then merged with the CNet files, which already contains charges information from CNet. "PRIM_unique_charge" will also be merged with the rest files and then row-binded to the file generated above.


### comb2 + PRIM_charges
```{r}
comb3 = merge(x=comb2, 
              y=PRIM_charges,
              by = c("PersonID", "InterviewID", "CaseID", "SeqCaseID"))
```

```{r}
a = comb3 %>% 
    group_by(PersonID, SeqCaseID, InterviewID, CaseID) %>% 
    summarise(count = n()) %>% 
    arrange(desc(count))
b = a[a$count>1, ]
sum(b$count - 1) == (nrow(comb3) - nrow(comb2))
```

__NOTE__: Extra records added. It is because each case could have multiple charges. Sanity check conducted above.

```{r}
colnames(comb3)
```

```{r}
comb3 = comb3 %>% 
    select(-County.y,
           -Division.y,
           -CaseNumberType.y) %>% 
    mutate(County = County.x,
           Divison = Division.x,
           CaseNumberType = CaseNumberType.x)
```



```{r}
PRIM_cases = comb2
PRIM_charges = comb3
PRIM_bond = Bonds
save(PRIM_cases, file = "PRIM_Cases.RData")
save(PRIM_charges, file = "PRIM_Charges.RData")
save(PRIM_bond, file = "PRIM_Bonds.RData")
```


