---
title: "Exploratory Analysis"
author: "Bin Han"
date: "April 27, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(data.table)
```


This file does simple exploratory data analysis of 12 text files.

```{r}
data_path = "../data/"
```

### KY1 -- defendants
```{r}
KY1 = as.data.frame(fread(paste0(data_path,'01_KY_Recidivism_Defendants.txt'), header = T))
```

__NOtice__:

Using data.table::fread to read the data is much faster and it can automatically get rid the of NA values. 


```{r}
names(KY1)
head(KY1)
length(unique(KY1$PersonID)) == nrow(KY1)
```

1. Each record has an unique PersonaID
2. No NA's (noticed when load the data using `fread`)


############################################################################################

### KY2 -- Interviews
```{r}
KY2 = as.data.frame(fread(paste0(data_path,'02_KY_Recidivism_Interviews.txt'), header = T))
```

```{r}
names(KY2)

length(unique(KY2$PersonId))
length(unique(KY2$InterviewId))

KY2 %>%
  group_by(PersonId) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

a = KY2[KY2$PersonId == 28566, ]
```

__Notice__:

Some personIDs appear more than once. 


### Arnold (Laura and Arnold Foundation) PSA: public safety assessment
- FTA: failure to appear
- NCA: new criminal activity
- NVCA: new violent criminal activity


#### FTA
```{r}
### Pick FTA Features
FTA = KY2 %>% 
  mutate(RA2_Has_Prior_Conviction = ifelse(RA2_Has_Prior_Misdemeanor=="2" & RA2_Has_Prior_Felony == "2", 0, 
                                    ifelse(RA2_Has_Prior_Misdemeanor == "1" |   RA2_Has_Prior_Felony=="1",1, NA)),
         RA2_New_Charge_While_Case_Pending = ifelse(RA2_New_Charge_While_Case_Pending == "1", 1, 
                                             ifelse(RA2_New_Charge_While_Case_Pending == "2", 0, NA)),
         RA2_How_Many_FTA_Last_2_Yrs = ifelse(RA2_How_Many_FTA_Last_2_Yrs == "1", 0,
                                       ifelse(RA2_How_Many_FTA_Last_2_Yrs == "2", 2,
                                       ifelse(RA2_How_Many_FTA_Last_2_Yrs == "3", 4, NA))),
         RA2_Has_Any_FTA_Older_2_Yrs = ifelse(RA2_Has_Any_FTA_Older_2_Yrs == "1", 1,
                                       ifelse(RA2_Has_Any_FTA_Older_2_Yrs == "2", 0, NA))) %>%
  select(RA2_New_Charge_While_Case_Pending, 
         RA2_Has_Prior_Conviction, 
         RA2_How_Many_FTA_Last_2_Yrs,
         RA2_Has_Any_FTA_Older_2_Yrs)

sum(FTA$RA2_New_Charge_While_Case_Pending == 1, na.rm = T)
sum(FTA$RA2_Has_Prior_Conviction == 1, na.rm = T)
sum(FTA$RA2_How_Many_FTA_Last_2_Yrs == 0, na.rm = T)
sum(FTA$RA2_Has_Any_FTA_Older_2_Yrs == 1, na.rm = T)

### Calculate Scores
FTA$Score = apply(FTA, 1, sum)


sum(is.na(FTA$Score))
sum(is.na(FTA$RA2_New_Charge_While_Case_Pending))
sum(is.na(FTA$RA2_Has_Prior_Conviction))
sum(is.na(FTA$RA2_How_Many_FTA_Last_2_Yrs))
sum(is.na(FTA$RA2_Has_Any_FTA_Older_2_Yrs))


## Check FTA score from Table with calculated score
test = data.frame(cbind(original = KY2$fta_risk_score_raw, 
             calculated = FTA$Score)) %>% 
  mutate(original = as.numeric(ifelse(original == "NULL", NA, as.character(original))))

sum(is.na(test$original))
sum(is.na(test$calculated))

sum(test$original == test$calculated, na.rm = T)
a =test[!(test$original == test$calculated), ]
```

__Wanring__:

1. Following Beau's code, I only used RA2, which is effective from 7/1/2013 until current. (It makes sense since from the code book, the raw FTA score is only valid staring 7/1/2013)
2. The way I calculated "Has Prior Felony or Misdemeanor" is different from Beau's method (His code has some logical problem, I think). Other features are the same. 
3. The value of "RA2_How_Many_FTA_Last_2_Yrs" from the documentation does not match with the way how FTA score is organized. FTA uses: 0 FTA, 1 FTA, and 2 or more FTA as cut point; documentation: 0 FTA, 1 or 2 FTA, and 3 or more FTA. 
4. The way Beau calculated FTA is simply adding all the numbers together. But according to the official FTA score calculation table, "RA2_How_Many_FTA_Last_2_Yrs" should be modified a little bit. For instance, 0 FTA - 0 point; 1 FTA - 2 points; 2 or more FTA - 4 points.
5. There FTA provided by the table has 13 more observations that are non-missing. Out of those non-missing calculated values (777198), only 2 are different.


#### NCA
```{r}
### Pick NCA Features
NCA = KY2 %>% 
  mutate(RA2_Age_At_Interview = ifelse(RA2_Age_At_Interview == "3", 0,
                                ifelse(RA2_Age_At_Interview == "NULL", NA, 2)),
         RA2_New_Charge_While_Case_Pending = ifelse(RA2_New_Charge_While_Case_Pending == "1", 3, 
                                             ifelse(RA2_New_Charge_While_Case_Pending == "2", 0, NA)),
         RA2_Has_Prior_Misdemeanor = ifelse(RA2_Has_Prior_Misdemeanor == "1", 1,
                                     ifelse(RA2_Has_Prior_Misdemeanor == "2", 0, NA)),
         RA2_Has_Prior_Felony = ifelse(RA2_Has_Prior_Felony == "1", 1,
                                ifelse(RA2_Has_Prior_Felony == "2", 0, NA)),
         RA2_Num_Prior_Violent = ifelse(RA2_Num_Prior_Violent == "1", 0,
                                 ifelse(RA2_Num_Prior_Violent == "2", 1,
                                 ifelse(RA2_Num_Prior_Violent == "3", 2, NA))),
         RA2_How_Many_FTA_Last_2_Yrs = ifelse(RA2_How_Many_FTA_Last_2_Yrs == "1", 0,
                                       ifelse(RA2_How_Many_FTA_Last_2_Yrs == "2", 1,
                                       ifelse(RA2_How_Many_FTA_Last_2_Yrs == "3", 2, NA))),
         RA2_Has_Prior_Incarceration = ifelse(RA2_Has_Prior_Incarceration == "1", 2,
                                       ifelse(RA2_Has_Prior_Incarceration == "2", 0, NA))) %>%
  select(RA2_Age_At_Interview, RA2_New_Charge_While_Case_Pending, RA2_Has_Prior_Felony,
         RA2_Has_Prior_Misdemeanor, RA2_Num_Prior_Violent, RA2_How_Many_FTA_Last_2_Yrs,
         RA2_Has_Prior_Incarceration)

## Calculate unweighted Score
NCA$Score = apply(NCA, 1, sum)

sum(is.na(NCA$Score))
sum(is.na(NCA$RA2_New_Charge_While_Case_Pending))
sum(is.na(NCA$RA2_Age_At_Interview))
sum(is.na(NCA$RA2_Has_Prior_Felony))
sum(is.na(NCA$RA2_Has_Prior_Misdemeanor))
sum(is.na(NCA$RA2_Num_Prior_Violent))
sum(is.na(NCA$RA2_Has_Prior_Incarceration))
sum(is.na(NCA$RA2_How_Many_FTA_Last_2_Yrs))

## Compare calculated NCA with NCA from the table
test = data.frame(cbind(original = KY2$nca_risk_score_raw,
             calculated = NCA$Score)) %>% 
  mutate(original = as.numeric(ifelse(original == "NULL", NA, as.character(original))))

sum(is.na(test$original))
sum(is.na(test$calculated))

sum(test$original == test$calculated, na.rm = T)
```

__Warning__:

1. Same issue as FTA: "RA2_How_Many_FTA_Last_2_Yrs" documented score does not match with the official calculation.
2. The calculated score has over 100,000 NA's more than the provided row NCA score. And the two variables barely have the same score for each individual.

#### NVCA: New Violent Criminal Activity
```{r}
### pick up NVCA features
NVCA = KY2 %>% 
  mutate(RA2_Current_Charge_Violent = ifelse(RA2_Current_Charge_Violent == "2", 0, 
                                      ifelse(RA2_Current_Charge_Violent == "1", 2, NA)),
         RA2_Violent_and_under_21 = ifelse(RA2_Violent_and_under_21 == "2", 0,
                                    ifelse(RA2_Violent_and_under_21 == "1", 1, NA)),
         RA2_New_Charge_While_Case_Pending = ifelse(RA2_New_Charge_While_Case_Pending == "2", 0,
                                             ifelse(RA2_New_Charge_While_Case_Pending == "1",1, NA)),
         RA2_Has_Prior_Conviction = ifelse(RA2_Has_Prior_Misdemeanor=="2" & RA2_Has_Prior_Felony == "2", 0, 
                                    ifelse(RA2_Has_Prior_Misdemeanor == "1" |   RA2_Has_Prior_Felony=="1", 1, NA)),
         RA2_Num_Prior_Violent = ifelse(RA2_Num_Prior_Violent == "1", 0,
                                 ifelse(RA2_Num_Prior_Violent == "2", 1,
                                 ifelse(RA2_Num_Prior_Violent == "3", 2, NA)))) %>% 
  select(RA2_Current_Charge_Violent, 
         RA2_Violent_and_under_21,
         RA2_New_Charge_While_Case_Pending,
         RA2_Has_Prior_Conviction,
         RA2_Num_Prior_Violent)

## NVCA unweighted score
NVCA$Score = apply(NVCA, 1, sum)

sum(is.na(NVCA$RA2_Current_Charge_Violent))
sum(is.na(NVCA$RA2_Violent_and_under_21))
sum(is.na(NVCA$RA2_New_Charge_While_Case_Pending))
sum(is.na(NVCA$RA2_Has_Prior_Conviction))
sum(is.na(NVCA$RA2_Num_Prior_Violent))
sum(is.na(NVCA$Score))

## Compare calculated NCA with NCA from the table
test = data.frame(cbind(original = KY2$pvf_risk_score_raw,
             calculated = NVCA$Score)) %>% 
  mutate(original = as.numeric(ifelse(original == "NULL", NA, as.character(original))))

sum(!is.na(test$original))
sum(!is.na(test$calculated))
sum(test$original == test$calculated, na.rm = T)
```

__Warning__:
1. Our calculated score has 777194 NA's, while the original score has 777213. 19 records gap.
But using all the non-NA calculated score to compare, we have 771127 same scores (6067 difference).

The question is: which one to use?

#### Combine all the scores
```{r}
Scores = data.frame(
  FTA = FTA$Score,
  NCA = NCA$Score,
  NVCA = NVCA$Score
)

Scores$all_na = apply(Scores,1,function(x) all(is.na(x)))
Scores$any_na = apply(Scores,1,function(x) any(is.na(x)))

df = Scores %>%
  filter(!any_na) %>%
  select(FTA,NCA,NVCA) %>%
  mutate_all(as.factor)

ggpairs(df)

```

__Summary__:

1. Compared with Compass data, we already have all the features and FTA, NCA, NVCA scores (even though need some clarification)
2. For riskSLIM data structure or format, we dont have features like "jail30", "p_drug", "p_traffic" etc. yet. 


#################################################################################
### KY3 - 
```{r}
KY3 = as.data.frame(fread(paste0(data_path,'03_KY_Recidivism_CNet_Cases.txt'), header = T))
```

```{r}
names(KY3)
```

```{r}
names(KY3)

### Check PersonID
length(unique(KY3$PersonID))
length(unique(KY3$SeqCaseID))


KY3 %>%
  group_by(PersonID)%>%
  summarize(count = n()) %>%
  arrange(desc(count))

a = KY3[KY3$PersonID == "19143", ]
```

__Warning__:

One PersonID could appear multiple times. But SeqCaseID is unique for each record. And multiple PRIM cases could link to the same SeqCaseID.


############################################################################################
### KY4 - PRIM Cases
```{r}
KY4 = as.data.frame(fread(paste0(data_path,'04_KY_Recidivism_PRIM_Cases.txt'), header = T))
```

```{r}
names(KY4)
```

```{r}
length(unique(KY4$PersonID))
length(unique(KY4$InterviewID))
length(unique(KY4$SeqCaseId))
length(unique(KY4$CaseID))
```

## same personID, same interviewID, same SeqCaseID, differnt CaseID
```{r}
KY4 %>% 
  group_by(PersonID, InterviewID, SeqCaseId) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

```{r}
a = KY4[KY4$PersonID == "409241", ]
b = KY4[KY4$PersonID == "97603" & KY4$SeqCaseId == "1440979", ]
c = KY4[KY4$PersonID == "358053", ]
d = KY4[KY4$PersonID == "365361" & KY4$SeqCaseId == "1466009", ]
e = KY4[KY4$PersonID == "128242" & KY4$SeqCaseId == "489665", ]
```

## same PersonID, same SeqCaseID, different interviewID, different CaseID
```{r}
KY4 %>% 
  group_by(PersonID, SeqCaseId) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

```{r}
f = KY4[KY4$PersonID == "152948" & KY4$SeqCaseId == "913837", ]
g = comb2[comb2$PersonID == "152948" & comb2$SeqCaseID == "913837", ]
```


```{r}

comb2 %>% 
  group_by(PersonID, SeqCaseID) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  .[1:5, ]


t = CNet_charges[CNet_charges$PersonID == "409581" & CNet_charges$SeqCaseID == "1715219", ]
h = comb2[comb2$PersonID == "409581" & comb2$SeqCaseID == "1715219", ]
```



## same PersonID, different other IDs
```{r}
d = KY4[KY4$PersonID == "152948", ]
```



__Warning__:

Each person could have more than one interview. CaseID is unique in KY4 as it repsents the case-level record in PRIM. SeqCaseID's are not unique. 


############################################################################################
### KY5 - PRIM Case_Cross_Reference
```{r}
KY5 = as.data.frame(fread(paste0(data_path,'05_KY_Recidivism_Case_Cross_Reference.txt'), header = T))
```

```{r}
names(KY5)
```

############################################################################################
### KY6 - CourtNet Charges
```{r}
KY6 = as.data.frame(fread(paste0(data_path,'06_KY_Recidivism_CNet_Charges.txt'), header = T))
```

```{r}
names(KY6)
head(KY6)

length(unique(KY6$SeqCaseID))
length(unique(KY6$PersonId))
```


############################################################################################
### KY7 - PRIM Charges
```{r}
KY7 = as.data.frame(fread(paste0(data_path,'07_KY_Recidivism_PRIM_Charges.txt'), header = T))
```

############################################################################################
### KY8 - Sentences
```{r}
KY8 = as.data.frame(fread(paste0(data_path,'08_KY_Recidivism_Sentences.txt'), header = T))
```

```{r}
names(KY8)
head(KY8)
```

############################################################################################
### KY9 - Bonds
```{r}
KY9 = as.data.frame(fread(paste0(data_path,'09_KY_Recidivism_Bonds.txt'), header = T))
```

```{r}
names(KY9)
```

```{r}
head(KY9)
```

```{r}
length(unique(KY9$PersonId))
length(unique(KY9$interviewID))
length(unique(KY9$CaseId))
length(unique(KY9$BondID))

```

```{r}
a = KY9 %>% select(interviewID, BondID, BondDate, IsReleased)
```


############################################################################################
### KY10 - Events
```{r}
KY10 = as.data.frame(fread(paste0(data_path,'10_KY_Recidivism_Events.txt'), header = T))
```

```{r}
names(KY10)
```

```{r}
head(KY10)
```

############################################################################################
### KY11 - FTA
```{r}
KY11 = as.data.frame(fread(paste0(data_path,'11_KY_Recidivism_FTA.txt'), header = T))
```

```{r}
head(KY11)
```

############################################################################################
### KY12 - Events
```{r}
KY12 = as.data.frame(fread(paste0(data_path,'12_KY_Recidivism_Supervision.txt'), header = T))
```


```{r}
write.csv(train, file = "train.csv")
write.csv(test, file = "test.csv")
```

